---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# `{serodynamics}`

<!-- badges: start -->

```{r badges, echo=FALSE, results='asis'}
# Use badger package to generate badges
cat(
  badger::badge_codecov("ucdavis/serodynamics", branch = "main"), "\n",
  "[![CodeFactor](https://www.codefactor.io/repository/github/ucdavis/serodynamics/badge)](https://www.codefactor.io/repository/github/ucdavis/serodynamics)\n",
  badger::badge_github_actions("ucdavis/serodynamics", "R-CMD-check.yaml"), "\n",
  badger::badge_cran_release("serodynamics", "orange"), "\n",
  badger::badge_lifecycle("experimental"), "\n"
)
```

<!-- badges: end -->

The goal of `{serodynamics}` is to implement methods for modeling longitudinal antibody responses to infection.

The package provides Bayesian MCMC modeling capabilities using either:

- **JAGS** (Just Another Gibbs Sampler) via `runjags` - the original implementation
- **Stan** via `cmdstanr` - a modern, efficient alternative (optional)

Both interfaces use the same data preparation and analysis workflow, allowing users to choose their preferred Bayesian modeling framework.

## Installation

You can install the development version of `{serodynamics}` from [GitHub](https://github.com/) with:

``` r
# install.packages("pak")
pak::pak("ucdavis/serodynamics")
```

### Stan Support (Optional)

To use Stan models (via `run_mod_stan()`), you'll also need to install `cmdstanr`:

``` r
# Install cmdstanr from r-universe
install.packages("cmdstanr", 
                 repos = c("https://stan-dev.r-universe.dev",
                          getOption("repos")))

# Then install CmdStan
cmdstanr::install_cmdstan()
```

See the [cmdstanr documentation](https://mc-stan.org/cmdstanr/) for more details.
