---
title: "Antibody dynamics for Cholera"
date: "`r format(Sys.time(), '%d %B, %Y')`"
author: "Samuel Schildhauer"
format: pdf
editor: visual
---

```{r echo=F, warning=F, message=F, include=F}
## Data preparation and running Jags

devtools::load_all()

#devtools::install_github("ucd-serg/serocalculator")
library(serocalculator)
library(tidyverse)
library(runjags)
library(rjags)
library(coda)
library(ggmcmc)
library(ggbeeswarm)
library(ggpubr)
library(rethinking)

#model file
file.mod <- here::here()  %>% fs::path("inst/extdata/model.jags.r")


#long data - TYPHOID 
# dL <-
# # the raw data is prepared and shared by jessica Seidman
#   read_csv(here::here()  %>% fs::path("inst/extdata/elisa_clean_2023-11-01.csv")) %>%
#  filter(surgical != 1 | is.na(surgical))  %>%
#   filter(Arm == "Prospective Cases" | Arm == "Retrospective Cases") %>%
#   mutate(Hospitalized = ifelse((recloc == "Inpatient Department" | admithosp_seap == "Yes"), "Yes", "No")) %>%
#   mutate(antigen_iso = paste(elisa_antigen, "_", elisa_antbdy_iso, sep="")) %>%
#   mutate(timeindays = ifelse(is.na(dayssincefeveronset), timesince0, dayssincefeveronset)) %>%
#   mutate(TimePeriod = factor(TimePeriod, levels = c("Baseline","First visit", "28 days", "3 months","6 months", "12 months", "18 months", "24 months", "Last visit"))) %>%
#   group_by(index_id, TimePeriod) %>% mutate(nVisits=n()) %>%
#   ungroup() %>%
#   select(index_id, Country, seapage, bldculres, Hospitalized, antigen_iso, result, TimePeriod,  postreinf, samplenum,  timeindays) %>%
#   rename(age = seapage) %>%
#   rename(visit_num = samplenum)  %>%
#   mutate(timeindays = ifelse(timeindays<0, 0, timeindays)) %>%
#   mutate(visit = ifelse(is.na(visit_num) & TimePeriod == "Baseline", 1, visit_num)) %>%
#   filter(!antigen_iso %in% c("YncE_IgG", "CdtB_IgA", "CdtB_IgG", "MP_IgA", "MP_IgG"))  %>%
#   droplevels()
#long data - CHOLERA
dL_cholera <-
  # read.csv("~/Library/CloudStorage/OneDrive-UniversityofCalifornia,Davis/Research/Cholera-longitudinal/data/cholera_data_compiled_050324.csv") %>%
  read_csv(here::here()  %>% fs::path("inst/extdata/cholera_data_compiled_050324.csv")) %>%
  group_by(index_id, antigen_iso) %>%                      # Group data by individual
  arrange(visit) %>%                          # Sort data by visit within each group
  mutate(visit_num = rank(visit, ties.method = "first")) %>%
  ungroup()

### Creating dummy variables for stratifications 
dL <- dL_cholera %>%
  mutate(Age_bin = ifelse(age_years <= 5 , "<6","6+"),
         Immunity_bin = ifelse(cohort=="Vaccinee" , "Vaccinee","Case"),
         Age_immune_combo = ifelse(cohort=="Vaccinee" & age_years <= 5, 
                                   "Vaccinated_5under",
                                   ifelse(cohort=="Vaccinee" & age_years > 5, 
                                          "Vaccinated_over5", 
                                          ifelse(cohort=="Case" & age_years > 5,
                                                 "Case_over5", 
                                                 ifelse(cohort=="Case" & age_years <= 5,
                                                 "Case_5under",NA)))))




```

Running full model, unstratified with the following inputs:

nchains \<- 4; \# nr of MC chains to run simultaneously

nadapt \<- 10000; \# nr of iterations for adaptation

nburnin \<- 100; \# nr of iterations to use for burn-in

nmc \<- 2500; \# nr of samples in posterior chains

niter \<- 5000; \# nr of iterations for posterior sample

nthin \<- round(niter/nmc); \# thinning needed to produce nmc from niter

```{r}

## Standardizing OD values 
# dL$result <- standardize(dL$result)

# Filtering down to two
# dL <- dL %>%
#   filter(antigen_iso %in% c("lpso139o_IgA"))

#set seed for reproducibility
set.seed(54321)
#subset data for checking
# dL_sub <- dL %>%
#   filter(index_id %in% sample(unique(index_id), 50))

#prepare data for modeline
longdata <- prep_data(dL)
priors <- prep_priors(max_antigens = longdata$n_antigen_isos)

#inputs for jags model
nchains <- 4;                # nr of MC chains to run simultaneously
nadapt  <- 10000;             # nr of iterations for adaptation
nburnin <- 100;            # nr of iterations to use for burn-in
nmc     <- 2500;             # nr of samples in posterior chains
niter   <- 5000;            # nr of iterations for posterior sample
nthin   <- round(niter/nmc); # thinning needed to produce nmc from niter

#pred.subj <- longdata$nsubj + 1;
#tomonitor <- c("par");
tomonitor <- c("y0", "y1", "t1", "alpha", "shape");

#This handles the seed to reproduce the results 
initsfunction <- function(chain){
  stopifnot(chain %in% (1:4)); # max 4 chains allowed...
  .RNG.seed <- (1:4)[chain];
  .RNG.name <- c("base::Wichmann-Hill","base::Marsaglia-Multicarry",
                 "base::Super-Duper","base::Mersenne-Twister")[chain];
  return(list(".RNG.seed"=.RNG.seed,".RNG.name"=.RNG.name));
}

jags.post <- run.jags(model=file.mod,data=c(longdata, priors),
                      inits=initsfunction,method="parallel",
                      adapt=nadapt,burnin=nburnin,thin=nthin,sample=nmc,
                      n.chains=nchains,
                      monitor=tomonitor,summarise=FALSE);
mcmc_list <- as.mcmc.list(jags.post)
mcmc_df <- ggs(mcmc_list)

wide_predpar_df <- mcmc_df %>%
  mutate(
    parameter = sub("^(\\w+)\\[.*", "\\1", Parameter),
    index_id = as.numeric(sub("^\\w+\\[(\\d+),.*", "\\1", Parameter)),
    antigen_iso = as.numeric(sub("^\\w+\\[\\d+,(\\d+).*", "\\1", Parameter))
  ) %>%
  mutate(
    index_id = factor(index_id, labels = attr(longdata, "ids")),
    antigen_iso = factor(antigen_iso, labels = attr(longdata, "antigens"))) %>%
 # mutate(value = exp(value)) %>%
 # mutate(value = ifelse(parameter == "r", value+1, value)) %>%
  ## only take the last subject (newperson)
  filter(index_id == "newperson") %>%
  select(-Parameter) %>%
  pivot_wider(names_from = "parameter", values_from="value") %>%
  rowwise() %>%
  #mutate(y1 = y0+y1) %>%
  droplevels() %>%
  ungroup() %>%
  rename(r = shape) 


#Now plot longitudinal antibody decay 
curve_params <- wide_predpar_df
  class(curve_params) =
  c("curve_params", class(curve_params))
  antigen_isos = unique(curve_params$antigen_iso)
  attr(curve_params, "antigen_isos") = antigen_isos

```

**Diagnostics of full model**

Note: Program is recommending additional adaptations – unsure how high to go. It sounds like this may be due to a complex model

**Summary**

```{r}

EffSize <- data.frame(effectiveSize(jags.post$mcmc))
EffSize <- rownames_to_column(EffSize)
EffSize <- EffSize %>%
  filter(grepl("159", rowname))
EffSize

```

**Traceplots**

```{r}
plot_jags_trace(data=jags.post,iso="lpso139o_IgA")
# posterior <- coda.samples(jags.post)
#   effectiveSize(jags.post)
```

**Number of effective samples**

```{r}
plot_jags_eff(data=jags.post,iso="lpso139o_IgA")

```

**Density plots**

```{r}
plot_jags_dens(data=jags.post,iso="lpso139o_IgA")
```

Running stratified data

```{r}

#Creating a function to run stratified data
Run_jags_fun <- function(data, strat) {
strat_list <<- unique(data[[strat]])

for (i in strat_list) {
dL_sub <- data %>%
  filter(data[[strat]]==i)
# dL_sub <- dL %>%
#   filter(Age_immune_combo=="Case_5under")

#set seed for reproducibility
set.seed(54321)
#subset data for checking
# dL_sub <- dL %>%
#   filter(index_id %in% sample(unique(index_id), 50))

#prepare data for modeline
longdata <- prep_data(dL_sub)
priors <- prep_priors(max_antigens = longdata$n_antigen_isos)

#inputs for jags model
nchains <- 4;                # nr of MC chains to run simultaneously
nadapt  <- 10000;             # nr of iterations for adaptation
nburnin <- 100;            # nr of iterations to use for burn-in
nmc     <- 2500;             # nr of samples in posterior chains
niter   <- 5000;            # nr of iterations for posterior sample
nthin   <- round(niter/nmc); # thinning needed to produce nmc from niter

#pred.subj <- longdata$nsubj + 1;
#tomonitor <- c("par");
tomonitor <- c("y0", "y1", "t1", "alpha", "shape");

#This handles the seed to reproduce the results 
initsfunction <- function(chain){
  stopifnot(chain %in% (1:4)); # max 4 chains allowed...
  .RNG.seed <- (1:4)[chain];
  .RNG.name <- c("base::Wichmann-Hill","base::Marsaglia-Multicarry",
                 "base::Super-Duper","base::Mersenne-Twister")[chain];
  return(list(".RNG.seed"=.RNG.seed,".RNG.name"=.RNG.name));
}

jags.post <- run.jags(model=file.mod,data=c(longdata, priors),
                      inits=initsfunction,method="parallel",
                      adapt=nadapt,burnin=nburnin,thin=nthin,sample=nmc,
                      n.chains=nchains,
                      monitor=tomonitor,summarise=FALSE);
mcmc_list <- as.mcmc.list(jags.post)
mcmc_df <- ggs(mcmc_list)

wide_predpar_df <- mcmc_df %>%
  mutate(
    parameter = sub("^(\\w+)\\[.*", "\\1", Parameter),
    index_id = as.numeric(sub("^\\w+\\[(\\d+),.*", "\\1", Parameter)),
    antigen_iso = as.numeric(sub("^\\w+\\[\\d+,(\\d+).*", "\\1", Parameter))
  ) %>%
  mutate(
    index_id = factor(index_id, labels = attr(longdata, "ids")),
    antigen_iso = factor(antigen_iso, labels = attr(longdata, "antigens"))) %>%
 # mutate(value = exp(value)) %>%
 # mutate(value = ifelse(parameter == "r", value+1, value)) %>%
  ## only take the last subject (newperson)
  filter(index_id == "newperson") %>%
  select(-Parameter) %>%
  pivot_wider(names_from = "parameter", values_from="value") %>%
  rowwise() %>%
  #mutate(y1 = y0+y1) %>%
  droplevels() %>%
  ungroup() %>%
  rename(r = shape) %>%
  mutate(stratification=i)

assign(paste0("wide_predpar_",i), data.frame(wide_predpar_df), envir=globalenv())

# wide_prepdar_df_shell <- rbind(wide_predpar_df_shell, wide_predpar_df)
}
}

## Running function 
Run_jags_fun(dL, "Age_immune_combo")

# Putting dataframes back together
wide_predpar_final <- rbind(wide_predpar_Case_5under, wide_predpar_Case_over5, wide_predpar_Vaccinated_5under, wide_predpar_Vaccinated_over5)

#Now plot longitudinal antibody decay 
curve_params <- wide_predpar_final
  class(curve_params) =
  c("curve_params", class(curve_params))
  antigen_isos = unique(curve_params$antigen_iso)
  attr(curve_params, "antigen_isos") = antigen_isos
```

Methods: Cholera data was run through jags simulation using the following settings: inputs for jags model:

nchains \<- 4; \# nr of MC chains to run simultaneously

nadapt \<- 10000; \# nr of iterations for adaptation

nburnin \<- 100; \# nr of iterations to use for burn-in

nmc \<- 2500; \# nr of samples in posterior chains

niter \<- 2500; \# nr of iterations for posterior sample

nthin \<- round(niter/nmc); \# thinning needed to produce nmc from niter

Data from the "New person" were subset and used for the following figures.

Parameter Values by Antigen Type

```{r echo=F, warning=F, message=F}
ggplot(curve_params, aes(
  x = as.factor(antigen_iso),
  y = alpha
)) +
  # geom_beeswarm(size = .2, alpha = .3, aes(color = antigen_iso), show.legend = F) +
  geom_boxplot(outlier.colour = NA, fill = NA) +
  scale_y_log10() +
  theme_linedraw() +
  labs(x = "antigen - isotype", title="Boxplot of parameter alpha by antigen type")+
  theme(axis.text.x = element_text(angle=45, hjust=1, vjust=1)) +
  stat_compare_means(label="p.signif", size=3, hide.ns = T) 

ggplot(curve_params, aes(
  x = as.factor(antigen_iso),
  y = r
)) +
  # geom_beeswarm(size = .2, alpha = .3, aes(color = antigen_iso), show.legend = F) +
  geom_boxplot(outlier.colour = NA, fill = NA) +
  scale_y_log10() +
  theme_linedraw() +
  labs(x = "antigen - isotype", title="Boxplot of parameter r by antigen type")+
  theme(axis.text.x = element_text(angle=45, hjust=1, vjust=1))+
  stat_compare_means(label="p.signif", size=3, hide.ns = T) 

ggplot(curve_params, aes(
  x = as.factor(antigen_iso),
  y = y1
)) +
  # geom_beeswarm(size = .2, alpha = .3, aes(color = antigen_iso), show.legend = F) +
  geom_boxplot(outlier.colour = NA, fill = NA) +
  scale_y_log10() +
  theme_linedraw() +
  labs(x = "antigen - isotype", title="Boxplot of parameter y1 by antigen type")+
  theme(axis.text.x = element_text(angle=45, hjust=1, vjust=1))+
  stat_compare_means(label="p.signif", size=3, hide.ns = T) 


ggplot(curve_params, aes(
  x = as.factor(antigen_iso),
  y = y0
)) +
  # geom_beeswarm(size = .2, alpha = .3, aes(color = antigen_iso), show.legend = F) +
  geom_boxplot(outlier.colour = NA, fill = NA) +
  scale_y_log10() +
  theme_linedraw() +
  labs(x = "antigen - isotype", title="Boxplot of parameter y0 by antigen type")+
  theme(axis.text.x = element_text(angle=45, hjust=1, vjust=1))+
  stat_compare_means(label="p.signif", size=3, hide.ns = T) 

ggplot(curve_params, aes(
  x = as.factor(antigen_iso),
  y = t1
)) +
  # geom_beeswarm(size = .2, alpha = .3, aes(color = antigen_iso), show.legend = F) +
  geom_boxplot(outlier.colour = NA, fill = NA) +
  scale_y_log10() +
  theme_linedraw() +
  labs(x = "antigen - isotype", title="Boxplot of parameter t1 by antigen type")+
  theme(axis.text.x = element_text(angle=45, hjust=1, vjust=1))+
  stat_compare_means(label="p.signif", size=3, hide.ns = T) 
```

```{r echo=F, warning=F, message=F, include=F}
### Prepping stratification by immunity
Run_jags_fun(dL, "Immunity_bin")
# Putting dataframes back together
wide_predpar_final_immunity <- rbind(wide_predpar_Vaccinee, wide_predpar_Case)
#Now plot longitudinal antibody decay 
curve_params_immunity <- wide_predpar_final_immunity
  class(curve_params) =
  c("curve_params", class(curve_params))
  antigen_isos = unique(curve_params$antigen_iso)
  attr(curve_params, "antigen_isos") = antigen_isos
```

Methods: Cholera data set was stratified by case vs vaccinee and $$\le 5$$ vs \>5 years of age and run through jags simulation using the following inputs: inputs for jags model nchains \<- 4; \# nr of MC chains to run simultaneously nadapt \<- 2000; \# nr of iterations for adaptation nburnin \<- 100; \# nr of iterations to use for burn-in nmc \<- 100; \# nr of samples in posterior chains niter \<- 1000; \# nr of iterations for posterior sample nthin \<- round(niter/nmc); \# thinning needed to produce nmc from niter

Antibody dynamic plots stratified by immunity status

```{r echo=F, warning=F, message=F}

jags_plot_function <- function(data=data, param, strat, label) {
ggplot(data, aes(
  x = as.factor(antigen_iso),
  y = .data[[param]]
)) +
  geom_boxplot(aes(color=.data[[strat]]),outlier.colour = NA, fill=NA) +
  # geom_beeswarm(size = .2, alpha = .3, show.legend = F) +
  scale_y_log10() +
  theme_classic() +
  labs(x = "antigen - isotype") +
  theme(axis.text.x = element_text(angle=45, hjust=1, vjust=1)) +
  labs(title=paste0("Median value for ", param," stratified by ",label))+
  stat_compare_means(aes(group=.data[[strat]]), label="p.signif", size=3, hide.ns = T) +
    scale_color_brewer(palette = "Paired")+
    guides(shape = guide_legend(override.aes = list(size = 1)))
}
jags_plot_function(curve_params_immunity, "y1", "stratification", "Immunity Type")
jags_plot_function(curve_params_immunity, "y0", "stratification", "Immunity Type")
jags_plot_function(curve_params_immunity, "t1", "stratification", "Immunity Type")
jags_plot_function(curve_params_immunity, "r", "stratification", "Immunity Type")
jags_plot_function(curve_params_immunity, "alpha", "stratification", "Immunity Type")


```

```{r echo=F, warning=F, message=F, include=F}
Run_jags_fun(dL, "Age_bin")
# Putting dataframes back together
wide_predpar_final_age <- rbind(`wide_predpar_6+`,`wide_predpar_<6`)
#Now plot longitudinal antibody decay 
curve_params_age <- wide_predpar_final_age
  class(curve_params) =
  c("curve_params", class(curve_params))
  antigen_isos = unique(curve_params$antigen_iso)
  attr(curve_params, "antigen_isos") = antigen_isos
```

Antibody dynamic plots stratified by age group

```{r echo=F, warning=F, message=F}
### Prepping stratification by age

jags_plot_function(curve_params_age, "y1", "stratification", "Age Category")
jags_plot_function(curve_params_age, "y0", "stratification", "Age Category")
jags_plot_function(curve_params_age, "t1", "stratification", "Age Category")
jags_plot_function(curve_params_age, "r", "stratification", "Age Category")
jags_plot_function(curve_params_age, "alpha", "stratification", "Age Category")
```

Antibody dynamic plots stratified by age group and immunity status

```{r echo=F, warning=F, message=F}
## Curve params for age AND immunity status

jags_plot_function(curve_params, "y1", "stratification", "Age and Immunity Status")
jags_plot_function(curve_params, "y0", "stratification", "Age and Immunity Status")
jags_plot_function(curve_params, "t1", "stratification", "Age and Immunity Status")
jags_plot_function(curve_params, "r", "stratification", "Age and Immunity Status")
jags_plot_function(curve_params, "alpha", "stratification", "Age and Immunity Status")
```

Antibody dynamic curves for each stratification

```{r echo=F, warning=F, message=F, include=F}
# 
# # autoplot(curve_params)
# autoplot(curve_params, log_x=T, n_curves=50)
# # graph.curve.params(curve_params) %>% print()
# ggplot(curve_params, aes(
#   x = as.factor(antigen_iso),
#   y = alpha
# )) +
#   geom_beeswarm(size = .2, alpha = .3, aes(color = antigen_iso), show.legend = F) +
#   geom_boxplot(outlier.colour = NA, fill = NA) +
#   scale_y_log10() +
#   theme_linedraw() +
#   labs(x = "antigen - isotype")+
#   theme(axis.text.x = element_text(angle=45, hjust=1, vjust=1))

```
