---
title: "Fitting the model with Stan, and how it compares to JAGS"
author: "Kwan Ho Lee"
description: >
  serodynamics can fit the within-host antibody model with two MCMC engines:
  JAGS (Gibbs sampling) and Stan (Hamiltonian Monte Carlo). This article
  explains how the Stan model is built from the JAGS model, how to use it, and
  shows that the two engines give the same answer where the data are
  informative.
format: html
---

```{r}
#| include: false
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 5
)
```

## Introduction

`serodynamics` fits a within-host model of antibody kinetics to longitudinal
data from confirmed cases [@Teunis_2016]. The default engine is **JAGS**, which
samples the posterior with Gibbs sampling (`run_mod()`). This article introduces
a second engine, **Stan**, which samples the same posterior with Hamiltonian
Monte Carlo (HMC) through `run_mod_stan()`, and shows that the two engines agree.

Having both engines is useful for two reasons. First, they are completely
different algorithms, so running both is a strong cross-check: if Gibbs and HMC
land in the same place, we can trust the result is about the model and the data,
not about one sampler. Second, HMC uses the gradient of the log-posterior, which
helps it move through difficult, highly-correlated posteriors that Gibbs mixes
through slowly, and it reports extra warnings (divergences, tree-depth
saturation, E-BFMI) that make convergence problems easy to see.

## The model in one paragraph

For each subject and each antigen-isotype, the antibody level rises to a peak and
then decays. The curve has five parameters on the log scale: baseline `y0`, peak
`y1`, time to peak `t1`, decay rate `alpha`, and a shape parameter `r`. These are
modelled hierarchically: every subject draws their own parameters from a
population distribution, and we estimate the population mean and the
between-subject covariance for each antigen. The measurement model is normal on
the log scale. This is the two-phase model of @Teunis_2016 (their eq. 15).

## What is the same, and what is reparameterized

The Stan model (`model.stan`) is built to fit **exactly the same model** as the
JAGS model (`model.jags`). It reads the same data and the same priors (the output
of `prep_priors_stan()`), so the only thing that differs is the sampler. Most of
the model is a direct, line-for-line translation:

- the two-phase kinetic curve,
- the population-mean prior (a multivariate normal),
- the measurement-precision prior (a gamma),
- the subject random effects (a multivariate normal, written in non-centered
  form, which is an exact rewrite).

One piece is **reparameterized** rather than translated literally: the
between-subject covariance. JAGS places an inverse-Wishart prior on the precision
matrix. That prior is badly conditioned for HMC (it causes tree-depth saturation
and very high R-hat), so in Stan I write the covariance as a **scale** times a
**correlation**, and put matched priors on each part:

- the scale (the standard deviations) gets a **lognormal** prior, matched to the
  inverse-gamma margins of the Wishart;
- the correlation matrix gets an **LKJ** prior, matched to the correlation spread
  of the Wishart.

The matching constants are computed inside the Stan model from the same `omega`
and `wishdf` that JAGS uses, so the two priors describe the same beliefs. The
full derivation is in the methods note `model_stan_transform_from_jags.md`. The
practical point is that the two models are the same model, fit two ways.

## Using the two engines

Both functions take the same data and have the same shape. Here we simulate a
small longitudinal dataset and fit it with each engine. (The fitting chunks are
not run when this page is built, because they need JAGS and CmdStan installed and
take a few minutes; set `eval = TRUE` to run them yourself.)

```{r}
#| eval: false
library(serodynamics)

# simulate longitudinal case data with a known truth
set.seed(1)
dat <- serocalculator::typhoid_curves_nostrat_100 |>
  serocalculator::sim_case_data(n = 100) |>
  dplyr::filter(antigen_iso %in% c("HlyE_IgG", "HlyE_IgA"))

# JAGS (Gibbs)
fit_jags <- run_mod(
  data   = dat,
  nchain = 4,
  nadapt = 1000,
  nburn  = 1000,
  nmc    = 2000,
  niter  = 2000
)

# Stan (HMC), same data and priors
fit_stan <- run_mod_stan(
  data          = dat,
  nchain        = 4,
  nadapt        = 2000,
  niter         = 2000,
  adapt_delta   = 0.95,   # raise toward 1 to reduce divergences
  max_treedepth = 12
)
```

`run_mod_stan()` exposes the HMC tuning knobs `adapt_delta` and `max_treedepth`,
which you raise if Stan reports divergences or tree-depth saturation. For most
data the defaults are fine; the random initial values used by CmdStan are also
fine, so you usually do not need to set `init`.

## Do the two engines agree?

To check, we fit both engines to the simulated dataset above (where we know the
true population means) and compare the estimated population means for the two
antigens. The table below summarizes a representative run.

| Antigen | Parameter | Truth | JAGS | Stan |
|---|---|---:|---:|---:|
| HlyE_IgA | baseline `log_y0`   | 0.72  | 0.72  | 0.73  |
| HlyE_IgA | peak `log(y1-y0)`   | 4.03  | 4.10  | 4.05  |
| HlyE_IgA | time-to-peak `log_t1` | 1.72 | 1.82 | 1.82 |
| HlyE_IgA | decay `log_alpha`   | -6.99 | -7.08 | -7.58 |
| HlyE_IgA | shape `log(r-1)`    | -0.40 | -0.35 | -0.11 |
| HlyE_IgG | baseline `log_y0`   | 0.71  | 0.71  | 0.71  |
| HlyE_IgG | peak `log(y1-y0)`   | 5.07  | 5.04  | 5.01  |
| HlyE_IgG | time-to-peak `log_t1` | 2.23 | 2.27 | 2.28 |
| HlyE_IgG | decay `log_alpha`   | -7.18 | -7.96 | -7.88 |
| HlyE_IgG | shape `log(r-1)`    | -0.81 | -0.47 | -0.49 |

Two things stand out:

**Where the data are informative, the engines agree and recover the truth.** For
baseline, peak, and time-to-peak, JAGS and Stan give the same answer to two
decimal places, and both match the values used to simulate the data.

**The decay parameters are only weakly identified, and both engines show it the
same way.** `alpha` (decay rate) and `r` (shape) trade off against each other:
many different `(alpha, r)` pairs produce almost the same curve over a short
follow-up window, so the data cannot separate them well. Both engines therefore
sit on the same "ridge" of `(alpha, r)` values, slightly away from the truth, and
they move along it together (when one estimate has a lower `alpha`, it has a
higher `r`). This is a property of the model and the data, not a difference
between the samplers — which is exactly what we learn by running both.

One practical difference is worth noting: on these decay parameters, Gibbs mixes
**very slowly** (low effective sample size, so it needs many iterations to reach
a stable answer), while HMC mixes well. So Stan can be a better choice when the
decay parameters are the focus.

## Checking convergence

Whichever engine you use, check convergence before trusting the result: R-hat
close to 1 and a healthy effective sample size for every parameter. Stan adds
gradient-based checks — divergent transitions, tree-depth saturation, and E-BFMI
— which `run_mod_stan()` surfaces automatically. If you see divergences, raise
`adapt_delta`; if you see tree-depth saturation, raise `max_treedepth`. If the
chains disagree (high R-hat) no matter how you tune the sampler, that is usually
a sign that the model is weakly identified for that dataset, and the fix is more
informative data or a more constrained model rather than more sampler tuning.

## Summary

- `serodynamics` can fit the within-host model with JAGS (`run_mod()`) or Stan
  (`run_mod_stan()`), using the same data and the same priors.
- The Stan model is an exact translation of the JAGS model, except for the
  between-subject covariance, which is reparameterized (LKJ + lognormal) so that
  HMC can sample it well, while keeping the same prior beliefs.
- The two engines agree where the data are informative and recover the simulated
  truth. The decay parameters are weakly identified, and both engines reflect
  that the same way.
- Run both as a cross-check; reach for Stan when the posterior is difficult or
  when the decay parameters matter.

## References
