---
title: "sim_case_data analysis"
author: "Samuel Schildhauer"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document:
    toc: true # table of content true
    toc_depth: 3  # upto three depths of headings (specified by #, ## and ###)
    toc_float: true
    #number_sections: true  ## if you want number sections at each table header
    #theme: united  # many options for theme, this one is my favorite.
    highlight: tango  # specifies the syntax highlighting style
---

### The goal
The goal of this vignette is to demonstrate the use of `sim_case_data()` and to
demonstrate at what sample size antibody kinetics can be accurately replicated.
To do so, we will model antibody kinetics from `nepal_sees` data. We will then 
simulate longitudinal case data using `sim_case_data()`. We will then determine
at what sample size the antibody kinetics, estimated with the entire `nepal_sees`
data set, can be reproduced using the simulated longitudinal data. 

### Loading packages
```{r  setup, error = F, warning = F, message = F}
library(dplyr)
library(tidyverse)
library(serocalculator)
library(gridExtra)
library(serodynamics)
knitr::opts_chunk$set(echo = TRUE)

```

  
### Plotting raw data  
Raw antibody responeses by individual  
```{r error = F, warning = F, message = F}

autoplot(serodynamics::nepal_sees)

```

### Modeling antibody kinetic parameters
First we will use run_serodynamics to model the antibody kinetics of the 
`nepal_sees` data set.  
```{r results = 'hide', error = F, warning = F, message = F}

#running data
nepal_sees_output <- run_serodynamics(data = serodynamics::nepal_sees,
                                      nchain = 4,
                                      nadapt = 100,
                                      nburn = 100,
                                      nmc = 500,
                                      niter = 2000,
                                      with_pop_params = TRUE)

```
  
### Antibody kinetic curves of modeled parameters  
Simulated antibody kinetic curves of the predictive distribution by number sampled  
```{r error = F, warning = F, message = F}
plot_serocurve(nepal_sees_output, log_y = TRUE)
```
  
Simulated antibody kinetic curves of the population-level distribution by number sampled  
```{r error = F, warning = F, message = F}
plot_serocurve(nepal_sees_output, log_y = TRUE, param_source = "population")
```

  
### Simulating case data  
Simulating case data using `as_case_data()` using the modeled antibody kinetic 
parameters as the curve parameters.  
```{r error = F, warning = F, message = F}
nepal_sees_prep <- nepal_sees_output %>%
  filter(Subject == "newperson") %>%
  spread(Parameter, value) %>%
  rename("r" = "shape",
         "antigen_iso" = "Iso_type") %>%
  group_by(antigen_iso) %>%
  mutate(iter = row_number()) 
# as_sr_params()


# Finding number of cases from original data set
n <- length(unique(nepal_sees$id))
antigen_isos <- unique(nepal_sees$antigen_iso)

# Simulating data
simulated_data <- sim_case_data(n = n,
                                   curve_params = nepal_sees_prep,
                                   antigen_isos = antigen_isos,
                                   followup_interval = 30,
                                   followup_variance = 30,
                                   noise_sd = 0.3)


```

### Plotting simulated data  
Raw longitudinal responses of simulated longitudinal data.  
```{r error = F, warning = F, message = F}
autoplot(simulated_data)
```
  
### Modeling antibody kinetic parameters on simulated longitudinal data
```{r results = 'hide', error = F, warning = F, message = F}

simulated_data <- simulated_data %>%
  as_case_data(id_var = "id",
               biomarker_var = "antigen_iso",
               value_var = "value",
               time_in_days = "timeindays")
n_select <- c("3", "5", "10", "30", "50", "100", "150", "188")
set.seed(15)
sero_list <- list()
for (i in n_select) {
  # for (s in 1:20) {
      for (s in 1:2) {
  sampled_n <- sample(unique(simulated_data$id), size = as.numeric(i))
  simulated_run <- simulated_data %>%
    filter(id %in% sampled_n) %>%
    select(id, timeindays, antigen_iso, value, visit_num)
    sero_run <- run_serodynamics(data = simulated_run,
                                        nchain = 4,
                                        nadapt = 100,
                                        nburn = 100,
                                        nmc = 500,
                                        niter = 2000,
                                        with_pop_params = TRUE)
  sero_run <- sero_run %>%
    mutate(strat = paste0(i, "_", s))
  sero_list[[paste0(i, "_", s)]] <- sero_run
  }
  }

saveRDS(sero_list, "~/projects/GitHub/serodynamics/inst/extdata/sim_serolist.rds")

```
  
### Looking at raw values of output
Look at predictive parameters
```{r}

post_pred_sim <- bind_rows(sero_list)
post_pred_summ <- post_pred_sim %>%
  group_by(Parameter, strat) %>%
  summarise(median = stats::median(value)) %>%
  mutate(n_samp = sub("_.*", "", strat)) %>%
  group_by(Parameter, n_samp) %>%
  summarise(median = stats::median(median))

# Need to also gather median from the original data

```

  
```{r}
post_pop_full <- list()
# Getting population parameters
for (i in names(sero_list)) {
  post_pop <- attributes(sero_list[[i]])$population_param %>%
    mutate(strat = i)
  post_pop_full <- bind_rows(post_pop_full, post_pop)
}

```

  
### Plotting the modeled serokinetic curve of simulated predictive distribution  
```{r error = F, warning = F, message = F}

# sero_list <- readRDS("~/projects/GitHub/serodynamics/inst/extdata/sim_serolist.rds")

# Creating plots 
plot_list <- list()
for (i in n_select) {
  sim_curve <- sero_list[[i]]
  plot_dat <- rbind(sim_curve,
                    nepal_sees_output %>%
                      mutate(Stratification = "full"))
  plot_list[[i]] <- plot_serocurve(plot_dat, log_y = T, xlim = c(0, 200)) +
    labs(title = paste0(i, " samples"))
}

```
  
Simulated antibody kinetic curves of predictive distribution by number sampled  
```{r error = F, warning = F, message = F, fig.height = 12}

grid.arrange(grobs = plot_list, ncol = 2)

```
  
### Plotting the modeled serokinetic curve of simulated population-level distribution  
```{r error = F, warning = F, message = F}

# sero_list <- readRDS("~/projects/GitHub/serodynamics/inst/extdata/sim_serolist.rds")

# Creating plots 
plot_list_pop <- list()
for (i in n_select) {
  sim_curve <- sero_list[[i]]
  sim_curve_pop <- attributes(sim_curve)$population_params %>%
    filter(Population_Parameter == "mu.par")  %>%
    mutate(Stratification = i)
  
  # Doing same for full model
  nepal_full_out <- attributes(nepal_sees_output)$population_params %>%
    filter(Population_Parameter == "mu.par") %>%
    mutate(Stratification = "full")
  
  # Putting the two together
  plot_dat <- rbind(sim_curve_pop,
                    nepal_full_out)
  attributes(plot_dat)$population_params <- plot_dat
  
  plot_list_pop[[i]] <- plot_serocurve(plot_dat, log_y = T, xlim = c(0, 200),
                                   param_source = "population") +
    labs(title = paste0(i, " samples"))
}

```
  
Simulated antibody kinetic curves of the population-level parameters by number sampled  
```{r error = F, warning = F, message = F, fig.height = 12}

grid.arrange(grobs = plot_list_pop, ncol = 2)

```


