---
title: "sim_case_data analysis"
author: "Samuel Schildhauer"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document:
    toc: true # table of content true
    toc_depth: 3  # upto three depths of headings (specified by #, ## and ###)
    toc_float: true
    #number_sections: true  ## if you want number sections at each table header
    #theme: united  # many options for theme, this one is my favorite.
    highlight: tango  # specifies the syntax highlighting style
    code_folding: hide
---

## The goal
The goal of this vignette is to demonstrate the use of `sim_case_data()` and to
demonstrate at what sample size antibody kinetics can be accurately replicated.
To do so, we will model antibody kinetics from `nepal_sees` data. We will then 
simulate longitudinal case data using `sim_case_data()`. We will then determine
at what sample size the antibody kinetics, estimated with the entire `nepal_sees`
data set, can be reproduced using the simulated longitudinal data. 

## Preparation
### Loading packages
```{r  setup, error = F, warning = F, message = F}
library(dplyr)
library(tidyverse)
library(serocalculator)
library(gridExtra)
library(serodynamics)
library(patchwork)
knitr::opts_chunk$set(echo = TRUE)

fig_num <- 1
```

  
### Plotting raw data  
Figure `r fig_num`: Raw antibody responeses by individual  
```{r error = F, warning = F, message = F}

autoplot(serodynamics::nepal_sees)

fig_num <- fig_num + 1
```

### Modeling antibody kinetic parameters
First we will use run_serodynamics to model the antibody kinetics of the 
`nepal_sees` data set.  
```{r results = 'hide', error = F, warning = F, message = F}

set.seed(15)
#running data
nepal_sees_output <- run_serodynamics(data = serodynamics::nepal_sees,
                                      nchain = 4,
                                      nadapt = 100,
                                      nburn = 100,
                                      nmc = 500,
                                      niter = 2000,
                                      with_pop_params = TRUE)

```
  
### Antibody kinetic curves of modeled parameters  
Figure `r fig_num`: Simulated antibody kinetic curves of the predictive distribution by number sampled  
```{r error = F, warning = F, message = F}
plot_serocurve(nepal_sees_output, log_y = TRUE)
fig_num <- fig_num + 1

```
  
Figure `r fig_num`: Simulated antibody kinetic curves of the population-level distribution by number sampled  
```{r error = F, warning = F, message = F}
plot_serocurve(nepal_sees_output, log_y = TRUE, param_source = "population")
fig_num <- fig_num + 1

```

## Simulated case data
### Simulating case data  
Simulating case data using `as_case_data()` using the modeled antibody kinetic 
parameters as the curve parameters.  
```{r error = F, warning = F, message = F}

nepal_sees_prep <- nepal_sees_output %>%
  filter(Subject == "newperson") %>%
  spread(Parameter, value) %>%
  rename("r" = "shape",
         "antigen_iso" = "Iso_type") %>%
  group_by(antigen_iso) %>%
  mutate(iter = row_number()) 
# as_sr_params()


# Finding number of cases from original data set
n <- length(unique(nepal_sees$id))
antigen_isos <- unique(nepal_sees$antigen_iso)


# Simulating data
simulated_data <- sim_case_data(n = n,
                                   curve_params = nepal_sees_prep,
                                   antigen_isos = antigen_isos,
                                   followup_interval = 20,
                                   followup_variance = 20,
                                   noise_sd = 0.4,
                                spacing = "log")


```
  
### Plotting simulated data  
Figure `r fig_num`: Raw longitudinal responses of simulated longitudinal data.  
```{r error = F, warning = F, message = F}
autoplot(simulated_data)
```
    
### Modeling antibody kinetic parameters on simulated longitudinal data  
We then take varying random samples of the simulated data to reconstruct the 
posterior distribution of original `nepal_sees` modeled data.  
```{r results = 'hide', error = F, warning = F, message = F, eval = F}

simulated_data <- simulated_data %>%
  as_case_data(id_var = "id",
               biomarker_var = "antigen_iso",
               value_var = "value",
               time_in_days = "timeindays")
n_select <- c("3", "5", "10", "30", "50", "100", "150", "188")
set.seed(15)
sero_list <- list()
for (i in n_select) {
  # for (s in 1:20) {
      for (s in 1:2) {
  sampled_n <- sample(unique(simulated_data$id), size = as.numeric(i))
  simulated_run <- simulated_data %>%
    filter(id %in% sampled_n) %>%
    select(id, timeindays, antigen_iso, value, visit_num)
    sero_run <- run_serodynamics(data = simulated_run,
                                        nchain = 4,
                                        nadapt = 100,
                                        nburn = 100,
                                        nmc = 500,
                                        niter = 2000,
                                        with_pop_params = TRUE)
  sero_run <- sero_run %>%
    mutate(sim = s, 
           n = i)
  sero_list[[paste0("n = ", i, ", sim = ", s)]] <- sero_run
  }
  }

saveRDS(sero_list, "~/projects/GitHub/serodynamics/inst/extdata/sero_list.rds")

```
  
### Parameter estimates of simulations  
We will then prepare the data of the modeled simulation outputs and combine them 
with the original modeled outputs.  
```{r  error = F, warning = F, message = F, fig.height=20}

sero_list <- readRDS("~/projects/GitHub/serodynamics/inst/extdata/sero_list.rds")

post_pred_sim <- bind_rows(sero_list)
post_pred_summ <- post_pred_sim %>%
  filter(Subject == "newperson") %>%
  group_by(Parameter, sim, n, Iso_type) %>%
  summarise(median = stats::median(value),
            quant2.5 = stats::quantile(value, 0.025),
            quant97.5 = stats::quantile(value, 0.975))%>%
  mutate(type = "Simulated posterior")

# Need to also gather median from the original data model
nepal_sees_summ <- nepal_sees_output %>%
  filter(Subject == "newperson") %>%
  group_by(Parameter, Iso_type) %>%
  summarise(median = stats::median(value),
            quant2.5 = stats::quantile(value, 0.025),
            quant97.5 = stats::quantile(value, 0.975)) %>%
  mutate(type = "Original posterior",
         sim = "Original")

nepal_sees_summ_bind <- data.frame(n = c(rep("3", 5),  rep("5", 5), rep("10", 5), 
                                         rep("30", 5), rep("50", 5), rep("100", 5), 
                                         rep("150", 5), rep("188", 5)), 
                                   Parameter = rep(c("alpha", "shape", "y1", "y0",
                                                     "t1"), 8))
nepal_sees_summ <- left_join(nepal_sees_summ, nepal_sees_summ_bind, by = "Parameter")

plot_pred_summ <- bind_rows(post_pred_summ %>%
                              mutate(sim = as.character(sim)), 
                            nepal_sees_summ) 

plot_pred_summ$n <- factor(plot_pred_summ$n, c("3", "5", "10", "30", "50", 
                                                         "100", "150", "188"))
plot_pred_summ$sim <- factor(plot_pred_summ$sim, rev(c("Original", 
                                                   as.character(unique(post_pred_summ$sim)))))

fig_num <- fig_num +1
```

We will then plot the results.  
Predictive parameters:  
Figure `r fig_num`: Posterior distribution of predictive parameters comparing modeled simulated to 
original data by number simulated for HlyE_IgA  
```{r  error = F, warning = F, message = F}

for (i in c("y0", "y1", "t1", "alpha", "shape")) {
print(
  ggplot(plot_pred_summ %>% filter(Parameter == i & Iso_type == "HlyE_IgA")) +
  geom_point(aes(x = median, y = sim, color = type)) +
  geom_linerange(aes(xmin = quant2.5, xmax = quant97.5, y = sim, color = type)) +
  scale_x_log10() +
  facet_wrap(~n, ncol = 4) +
   theme_bw() +
  labs(x = "Posterior distribution", y = "Simulation",
       title = paste0("Posterior predictive distribution for HlyE_IgA/", i, 
                      " \nby number sampled and simulation")) +
  scale_color_manual(values = c("skyblue1", "dodgerblue4")) +
    theme(legend.position = "top")
)
}
fig_num <- fig_num +1

```
  
Figure `r fig_num`: Posterior distribution of predictive parameters comparing modeled simulated to 
original data by number simulated for HlyE_IgG  
```{r  error = F, warning = F, message = F}

for (i in c("y0", "y1", "t1", "alpha", "shape")) {
print(
ggplot(plot_pred_summ %>% filter(Parameter == i & Iso_type == "HlyE_IgG")) +
  geom_point(aes(x = median, y = sim, color = type)) +
  geom_linerange(aes(xmin = quant2.5, xmax = quant97.5, y = sim, color = type)) +
  scale_x_log10() +
  facet_wrap(~n, ncol = 4) +
   theme_bw() +
  labs(x = "Posterior distribution", y = "Simulation",
       title = paste0("Posterior predictive distribution for HlyE_IgA/", i, 
                      " \\by number sampled and simulation")) +
  scale_color_manual(values = c("skyblue1", "dodgerblue4"))+
    theme(legend.position = "top")
)
}

fig_num <- fig_num +1

```

Population Parameters:  
Data preparation
```{r  error = F, warning = F, message = F}

# Extracting population parameters
pop_data <- list()
for (i in names(sero_list)) {
  pop_atts <- attributes(sero_list[[i]])$population_params
  n <- str_extract(i, "(?<=n = )\\d+")
  sim <- str_extract(i, "(?<=sim = )\\d+")
  pop_atts <- pop_atts %>%
    mutate(n =  n,
           sim = sim)
  pop_data <- dplyr::bind_rows(pop_data, pop_atts)
}

post_pop_summ <- pop_data %>%
  filter(Population_Parameter == "mu.par") %>%
  group_by(Parameter, sim, n, Iso_type) %>%
  summarise(median = stats::median(value),
            quant2.5 = stats::quantile(value, 0.025),
            quant97.5 = stats::quantile(value, 0.975))%>%
  mutate(type = "Simulated posterior")

# Need to also gather median from the original data model
nepal_pop_summ <- attributes(nepal_sees_output)$population_params %>%
  filter(Population_Parameter == "mu.par") %>%
  group_by(Parameter, Iso_type) %>%
  summarise(median = stats::median(value),
            quant2.5 = stats::quantile(value, 0.025),
            quant97.5 = stats::quantile(value, 0.975)) %>%
  mutate(type = "Original posterior",
         sim = "Original")

nepal_pop_summ_bind <- data.frame(n = c(rep("3", 5),  rep("5", 5), rep("10", 5), 
                                         rep("30", 5), rep("50", 5), rep("100", 5), 
                                         rep("150", 5), rep("188", 5)), 
                                   Parameter = rep(c("log(alpha)", "log(shape - 1)", "log(y1 - y0)", "log(y0)",
                                                     "log(t1)"), 8))
nepal_pop_summ <- left_join(nepal_pop_summ, nepal_pop_summ_bind, by = "Parameter")

plot_pop_summ <- bind_rows(post_pop_summ, nepal_pop_summ) 

plot_pop_summ$n <- factor(plot_pop_summ$n, c("3", "5", "10", "30", "50", 
                                                         "100", "150", "188"))
plot_pop_summ$sim <- factor(plot_pop_summ$sim, rev(c("Original", 
                                                   as.character(unique(post_pop_summ$sim)))))

```
  
Figure `r fig_num`: Posterior distribution of population parameters comparing modeled simulated to 
original data by number simulated for HlyE_IgA  
```{r  error = F, warning = F, message = F}

for (i in c("log(alpha)", "log(shape - 1)", "log(y1 - y0)", "log(y0)",
                                                     "log(t1)")) {
print(
ggplot(plot_pop_summ %>% filter(Parameter == i &  Iso_type == "HlyE_IgA")) +
  geom_point(aes(x = median, y = sim, color = type)) +
  geom_linerange(aes(xmin = quant2.5, xmax = quant97.5, y = sim, color = type)) +
  # scale_x_log10() +
  facet_wrap(~n, scales = "free_x", ncol = 4) +
   theme_bw() +
  labs(x = "Posterior distribution", y = "Simulation",
       title = paste0("Posterior population distribution for HlyE_IgA/", i, 
                      " \nby number sampled and simulation")) +
  scale_color_manual(values = c("skyblue1", "dodgerblue4")) +
    theme(legend.position = "top")
)
}
fig_num <- fig_num +1


```

  
Figure `r fig_num`: Posterior distribution of population parameters comparing modeled simulated to 
original data by number simulated for HlyE_IgG   
```{r  error = F, warning = F, message = F}

for (i in c("log(alpha)", "log(shape - 1)", "log(y1 - y0)", "log(y0)",
                                                     "log(t1)")) {
print(
ggplot(plot_pop_summ %>% filter(Parameter == i &  Iso_type == "HlyE_IgG")) +
  geom_point(aes(x = median, y = sim, color = type)) +
  geom_linerange(aes(xmin = quant2.5, xmax = quant97.5, y = sim, color = type)) +
  # scale_x_log10() +
  facet_wrap(~n, scales = "free_x") +
   theme_bw() +
  labs(x = "Posterior distribution", y = "Simulation",
       title = paste0("Posterior population distribution for HlyE_IgA/", i, 
                      " \nby number sampled and simulation")) +
  scale_color_manual(values = c("skyblue1", "dodgerblue4"))
)
}
fig_num <- fig_num +1

```


### RMSE analysis  
We will now calculate monte carlo bias comparing the median, 2.5% and 97.5% 
population-level posterior estimates for each parameter across samples taken. 
The estimates follow:  
$$
\mathrm{RMSE}_N =
\sqrt{
\frac{1}{S}
\sum_{s=1}^{S}
\left(
\hat{q}_{s,N} - q_{\mathrm{original}}
\right)^2
}
$$

Data preparation:
```{r error = F, warning = F, fig.height=15}

sim_post <- plot_pop_summ %>%
  filter(type == "Simulated posterior")

original_post <- plot_pop_summ %>%
  filter(type == "Original posterior") %>%
  dplyr::ungroup() %>%
  select(Parameter, Iso_type, median, quant2.5, quant97.5, type)

comparison <- sim_post %>%
  left_join(
    original_post %>%
      select(
        Parameter,
        Iso_type,
        original_lower = quant2.5,
        original_median = median,
        original_upper = quant97.5
      ),
    by = c("Parameter", "Iso_type")
  )

comparison <- comparison %>%
  mutate(
    error_lower = quant2.5 - original_lower,
    error_median = median - original_median,
    error_upper = quant97.5 - original_upper
  )

mc_bias <- comparison %>%
  group_by(n, Parameter, Iso_type) %>%
  summarise(
    bias_lower = mean(error_lower, na.rm = TRUE),
    bias_median = mean(error_median, na.rm = TRUE),
    bias_upper = mean(error_upper, na.rm = TRUE),
    .groups = "drop"
  )
mc_rmse <- comparison %>%
  group_by(n, Parameter, Iso_type) %>%
  summarise(
    rmse_lower = sqrt(mean(error_lower^2, na.rm = TRUE)),
    rmse_median = sqrt(mean(error_median^2, na.rm = TRUE)),
    rmse_upper = sqrt(mean(error_upper^2, na.rm = TRUE)),
    .groups = "drop"
  )

mc_rmse_long <- comparison %>%
  group_by(n, Parameter, Iso_type) %>%
  summarise(
    lower = sqrt(mean(error_lower^2, na.rm = TRUE)),
    median = sqrt(mean(error_median^2, na.rm = TRUE)),
    upper = sqrt(mean(error_upper^2, na.rm = TRUE)),
    .groups = "drop"
  ) %>%
  tidyr::pivot_longer(
    cols = c(lower, median, upper),
    names_to = "posterior_quantile",
    values_to = "RMSE"
  )


```

Figure `r fig_num`: Monte Carlo RMSE of median, 2.5% and 97.5% posterior distribution 
by number of simulated cases used for modeling and antigen isotype  
```{r error = F, warning = F, fig.height=12}

ggplot(
  mc_rmse_long,
  aes(x = n, y = RMSE, linetype = posterior_quantile,
      group = posterior_quantile)) +
  geom_line() +
  geom_point(size = 2) +
  facet_grid(
    Parameter ~ Iso_type,
    scales = "free_y"
  ) +
  theme_bw() +
  labs(
    x = "Number of subjects",
    y = "Monte Carlo RMSE",
    linetype = "Posterior summary"
  )+
    theme(legend.position = "top")

fig_num <- fig_num +1

```


### Modeled serokinetic curves  
Figure `r fig_num`: Serokinetic curves of simulated and original predictive 
posterior curves by number of simulations sampled and antigen isotype  
```{r error = F, warning = F, message = F}

# Need to also gather median from the original data model
nepal_pred_seroplot <- left_join(nepal_sees_output, nepal_sees_summ_bind,
                                by = "Parameter")

plot_pred_curves <- bind_rows(post_pred_sim, nepal_pred_seroplot) %>%
  mutate(Stratification = sim) %>%
  mutate(Stratification = ifelse(is.na(Stratification), "Original", 
                                 Stratification))
  # mutate(Stratification = paste0(Stratification, "_", n))


plot_list <- list()
for (i in unique(plot_pred_curves$n)) {
strat_levels <- unique(na.omit(plot_pred_curves$Stratification))
color_values <- setNames(
  ifelse(strat_levels == "Original", "red", "blue"),
  strat_levels
)
## Original data layer
plot_pred_curveplot <- plot_pred_curves %>%
  filter(n == i)
p_original <- plot_serocurve(
  plot_pred_curveplot %>% 
    filter(Stratification == "Original"),
  log_y = TRUE,
  show_ci = TRUE
) 
  # Make Original curve prominent
  p_original$layers <- lapply(
    p_original$layers,
    function(layer) {
      if (inherits(layer$geom, "GeomLine")) {
        layer$aes_params$color <- "red"
        layer$aes_params$alpha <- 1
        layer$aes_params$linewidth <- 1.2
      }
      layer
    }
  )
  
## Simulation layer
p_sim <- plot_serocurve(
  plot_pred_curveplot %>% 
    filter(Stratification != "Original"),
  log_y = TRUE,
  show_ci = FALSE
)
  p_sim$layers <- lapply(
    p_sim$layers,
    function(layer) {
      if (inherits(layer$geom, "GeomLine")) {
        layer$aes_params$color <- "blue"
        layer$aes_params$alpha <- 0.4
        layer$aes_params$linewidth <- 0.5
      }
      layer
    }
  )
  
plot_list[[i]] <- p_sim +
  p_original$layers +
  scale_color_manual(
    values = color_values,
    breaks = c(
      strat_levels[strat_levels != "Original"][1],
      "Original"
    ),
    labels = c("Simulation", "Original")
  ) +
  guides(
    color = guide_legend(
      title = NULL,
      override.aes = list(
        alpha = 1,
        linewidth = 1
      )
    ),
    fill = "none"
  ) +
  labs(title = paste0("n = ", i)) 
}
wrap_plots(plot_list, ncol = 2)  +
  plot_layout(guides = "collect") &
  theme(
    legend.position = "bottom"
  )

  
```


Figure `r fig_num`: Serokinetic curves of simulated and original population level 
posterior curves by number of simulations sampled and antigen isotype  
```{r error = F, warning = F, message = F}

# Need to also gather median from the original data model
nepal_pop_seroplot <- attributes(nepal_sees_output)$population_params 
nepal_pop_seroplot <- left_join(nepal_pop_seroplot, nepal_pop_summ_bind,
                                by = "Parameter")

plot_pop_curves <- bind_rows(pop_data, nepal_pop_seroplot) %>%
  mutate(Stratification = sim) %>%
  mutate(Stratification = ifelse(is.na(Stratification), "Original", 
                                 Stratification)) %>%
  filter(is.na(n)==F)

plot_list <- list()
for (i in unique(plot_pop_curves$n)) {
strat_levels <- unique(na.omit(plot_pop_curves$Stratification))
color_values <- setNames(
  ifelse(strat_levels == "Original", "red", "blue"),
  strat_levels
)
# Setting up data first
plot_pop_curvesplot <- plot_pop_curves %>%
  filter(n == i)
attr(plot_pop_curvesplot, "population_params") <- plot_pop_curvesplot

## Original data layer
p_original <- plot_serocurve(
  plot_pop_curvesplot %>% 
    filter(Stratification == "Original"),
  log_y = TRUE,
  show_ci = TRUE,
  param_source = "population"
) 
  # Make Original curve prominent
  p_original$layers <- lapply(
    p_original$layers,
    function(layer) {
      if (inherits(layer$geom, "GeomLine")) {
        layer$aes_params$color <- "red"
        layer$aes_params$alpha <- 1
        layer$aes_params$linewidth <- 1.2
      }
      layer
    }
  )
  
## Simulation layer
p_sim <- plot_serocurve(
  plot_pop_curvesplot %>% 
    filter(Stratification != "Original"),
  log_y = TRUE,
  show_ci = FALSE,
  param_source = "population"
)
  p_sim$layers <- lapply(
    p_sim$layers,
    function(layer) {
      if (inherits(layer$geom, "GeomLine")) {
        layer$aes_params$color <- "blue"
        layer$aes_params$alpha <- 0.4
        layer$aes_params$linewidth <- 0.5
      }
      layer
    }
  )
  
plot_list[[i]] <- p_sim +
  p_original$layers +
  scale_color_manual(
    values = color_values,
    breaks = c(
      strat_levels[strat_levels != "Original"][1],
      "Original"
    ),
    labels = c("Simulation", "Original")
  ) +
  guides(
    color = guide_legend(
      title = NULL,
      override.aes = list(
        alpha = 1,
        linewidth = 1
      )
    ),
    fill = "none"
  ) +
  labs(title = paste0("n = ", i)) 
}
wrap_plots(plot_list, ncol = 2)  +
  plot_layout(guides = "collect") &
  theme(
    legend.position = "bottom"
  )

  
```
  

### Rhat values by number of samples
Rhat values by number of samples for predictive parameters
```{r}

sim_pred_rhat <- post_pred_sim %>%
  filter(Subject == "newperson")
# Calculate Rhat
sim_pred_rhat <- sim_pred_rhat %>%
  group_by(Parameter, Iso_type, Stratification, sim, n) %>%
  group_modify(~{
    mat <- .x %>%
      arrange(Iteration, Chain) %>%
      pivot_wider(
        id_cols = Iteration,
        names_from = Chain,
        values_from = value
      ) %>%
      select(-Iteration) %>%
      as.matrix()
    tibble(Rhat = posterior::rhat(mat))
  }) %>%
  group_by(Parameter, Iso_type, n) %>%
  summarise(median = median(Rhat),
            lower = quantile(Rhat, 0.025),
            upper = quantile(Rhat, 0.975))

# sim_pred_rhat$n <- factor(sim_pred_rhat$n, c("3", "5", "10", "30", "50", 
#                                                          "100", "150", "188"))

ggplot(sim_pred_rhat, aes(x = as.numeric(n), y = median)) +
  geom_line(aes(x = as.numeric(n), y = median), stat = "identity") +
  geom_point(size = 2) +
  geom_linerange(aes(x = as.numeric(n), y = median, ymax = upper, ymin = lower)) +
  facet_grid(
    Parameter ~ Iso_type,
    scales = "free_y"
  ) +
  theme_bw() +
  labs(x = "Number of samples from posterior",
    y = "Rhat value"
  )+
    theme(legend.position = "top")

```


Rhat values by number of samples for population parameters
```{r error = F, warning = F}

sim_pop_rhat <- pop_data %>%
  filter(Population_Parameter == "mu.par")
# Calculate Rhat
sim_pop_rhat <- sim_pop_rhat %>%
  group_by(Parameter, Iso_type, Stratification, sim, n) %>%
  group_modify(~{
    mat <- .x %>%
      arrange(Iteration, Chain) %>%
      pivot_wider(
        id_cols = Iteration,
        names_from = Chain,
        values_from = value
      ) %>%
      select(-Iteration) %>%
      as.matrix()
    tibble(Rhat = posterior::rhat(mat))
  }) %>%
  group_by(Parameter, Iso_type, n) %>%
  summarise(median = median(Rhat),
            lower = quantile(Rhat, 0.025),
            upper = quantile(Rhat, 0.975))

# sim_pred_rhat$n <- factor(sim_pred_rhat$n, c("3", "5", "10", "30", "50", 
#                                                          "100", "150", "188"))

ggplot(sim_pop_rhat, aes(x = as.numeric(n), y = median)) +
  geom_line(aes(x = as.numeric(n), y = median), stat = "identity") +
  geom_point(size = 2) +
  geom_linerange(aes(x = as.numeric(n), y = median, ymax = upper, ymin = lower)) +
  facet_grid(
    Parameter ~ Iso_type,
    scales = "free_y"
  ) +
  theme_bw() +
  labs(x = "Number of samples from posterior",
    y = "Rhat value"
  )+
    theme(legend.position = "top")

```


## Sampling original data  
### Modeling antibody kinetic parameters on simulated longitudinal data  
We then take varying random samples of the simulated data to reconstruct the 
posterior distribution of original `nepal_sees` modeled data.  
```{r results = 'hide', error = F, warning = F, message = F, eval = F}

n_select <- c("3", "5", "10", "30", "50", "100", "150", "188")
set.seed(15)
nep_sero_list <- list()
for (i in n_select) {
  # for (s in 1:20) {
      for (s in 1:2) {
  sampled_n <- sample(unique(nepal_sees$id), size = as.numeric(i))
  simulated_run <- nepal_sees %>%
    filter(id %in% sampled_n) 
    sero_run <- run_serodynamics(data = simulated_run,
                                        nchain = 4,
                                        nadapt = 100,
                                        nburn = 100,
                                        nmc = 500,
                                        niter = 2000,
                                        with_pop_params = TRUE)
  sero_run <- sero_run %>%
    mutate(sim = s, 
           n = i)
  nep_sero_list[[paste0("n = ", i, ", sim = ", s)]] <- sero_run
  }
  }

saveRDS(sero_list, "~/projects/GitHub/serodynamics/inst/extdata/nep_serolist.rds")

```
  
### Parameter estimates of simulations  
We will then prepare the data of the modeled simulation outputs and combine them 
with the original modeled outputs.  
```{r  error = F, warning = F, message = F, fig.height=20}

nep_sero_list <- readRDS("~/projects/GitHub/serodynamics/inst/extdata/nep_serolist.rds")

neppost_pred_sim <- bind_rows(nep_sero_list)
neppost_pred_summ <- neppost_pred_sim %>%
  filter(Subject == "newperson") %>%
  group_by(Parameter, sim, n, Iso_type) %>%
  summarise(median = stats::median(value),
            quant2.5 = stats::quantile(value, 0.025),
            quant97.5 = stats::quantile(value, 0.975))%>%
  mutate(type = "Simulated posterior")

nepplot_pred_summ <- bind_rows(neppost_pred_summ %>%
                              mutate(sim = as.character(sim)), 
                            nepal_sees_summ) 

nepplot_pred_summ$n <- factor(nepplot_pred_summ$n, c("3", "5", "10", "30", "50", 
                                                         "100", "150", "188"))
nepplot_pred_summ$sim <- factor(nepplot_pred_summ$sim, rev(c("Original", 
                                                   as.character(unique(neppost_pred_summ$sim)))))

fig_num <- fig_num +1
```

We will then plot the results.  
Predictive parameters:  
Figure `r fig_num`: Posterior distribution of predictive parameters comparing varying samples of 
original data for HlyE_IgA  
```{r  error = F, warning = F, message = F}

for (i in c("y0", "y1", "t1", "alpha", "shape")) {
print(
  ggplot(nepplot_pred_summ %>% filter(Parameter == i & Iso_type == "HlyE_IgA")) +
  geom_point(aes(x = median, y = sim, color = type)) +
  geom_linerange(aes(xmin = quant2.5, xmax = quant97.5, y = sim, color = type)) +
  scale_x_log10() +
  facet_wrap(~n, ncol = 4) +
   theme_bw() +
  labs(x = "Posterior distribution", y = "Simulation",
       title = paste0("Posterior predictive distribution for HlyE_IgA/", i, 
                      " \nby number sampled and simulation")) +
  scale_color_manual(values = c("skyblue1", "dodgerblue4")) +
    theme(legend.position = "top")
)
}
fig_num <- fig_num +1

```
  
Figure `r fig_num`: Posterior distribution of predictive parameters comparing modeled simulated to 
original data by number simulated for HlyE_IgG  
```{r  error = F, warning = F, message = F}

for (i in c("y0", "y1", "t1", "alpha", "shape")) {
print(
ggplot(plot_pred_summ %>% filter(Parameter == i & Iso_type == "HlyE_IgG")) +
  geom_point(aes(x = median, y = sim, color = type)) +
  geom_linerange(aes(xmin = quant2.5, xmax = quant97.5, y = sim, color = type)) +
  scale_x_log10() +
  facet_wrap(~n, ncol = 4) +
   theme_bw() +
  labs(x = "Posterior distribution", y = "Simulation",
       title = paste0("Posterior predictive distribution for HlyE_IgA/", i, 
                      " \\by number sampled and simulation")) +
  scale_color_manual(values = c("skyblue1", "dodgerblue4"))+
    theme(legend.position = "top")
)
}

fig_num <- fig_num +1

```

Population Parameters:  
Data preparation
```{r  error = F, warning = F, message = F}

# Extracting population parameters
nep_pop_data <- list()
for (i in names(nep_sero_list)) {
  pop_atts <- attributes(nep_sero_list[[i]])$population_params
  n <- str_extract(i, "(?<=n = )\\d+")
  sim <- str_extract(i, "(?<=sim = )\\d+")
  pop_atts <- pop_atts %>%
    mutate(n =  n,
           sim = sim)
  nep_pop_data <- dplyr::bind_rows(nep_pop_data, pop_atts)
}

neppost_pop_summ <- nep_pop_data %>%
  filter(Population_Parameter == "mu.par") %>%
  group_by(Parameter, sim, n, Iso_type) %>%
  summarise(median = stats::median(value),
            quant2.5 = stats::quantile(value, 0.025),
            quant97.5 = stats::quantile(value, 0.975))%>%
  mutate(type = "Simulated posterior")

# Adding in nepal data
nepplot_pop_summ <- bind_rows(neppost_pop_summ, nepal_pop_summ) 

nepplot_pop_summ$n <- factor(nepplot_pop_summ$n, c("3", "5", "10", "30", "50", 
                                                         "100", "150", "188"))
nepplot_pop_summ$sim <- factor(nepplot_pop_summ$sim, rev(c("Original", 
                                                   as.character(unique(neppost_pop_summ$sim)))))

```
  
Figure `r fig_num`: Posterior distribution of predictive parameters comparing modeled simulated to 
original data by number simulated for HlyE_IgA  
```{r  error = F, warning = F, message = F}

for (i in c("log(alpha)", "log(shape - 1)", "log(y1 - y0)", "log(y0)",
                                                     "log(t1)")) {
print(
ggplot(nepplot_pop_summ %>% filter(Parameter == i &  Iso_type == "HlyE_IgA")) +
  geom_point(aes(x = median, y = sim, color = type)) +
  geom_linerange(aes(xmin = quant2.5, xmax = quant97.5, y = sim, color = type)) +
  # scale_x_log10() +
  facet_wrap(~n, ncol = 4) +
   theme_bw() +
  labs(x = "Posterior distribution", y = "Simulation",
       title = paste0("Posterior population distribution for HlyE_IgA/", i, 
                      " \nby number sampled and simulation")) +
  scale_color_manual(values = c("skyblue1", "dodgerblue4")) +
    theme(legend.position = "top")
)
}
fig_num <- fig_num +1


```

  
Figure `r fig_num`: Posterior distribution of predictive parameters comparing modeled simulated to 
original data by number simulated for HlyE_IgG   
```{r  error = F, warning = F, message = F}

for (i in c("log(alpha)", "log(shape - 1)", "log(y1 - y0)", "log(y0)",
                                                     "log(t1)")) {
print(
ggplot(plot_pop_summ %>% filter(Parameter == i &  Iso_type == "HlyE_IgG")) +
  geom_point(aes(x = median, y = sim, color = type)) +
  geom_linerange(aes(xmin = quant2.5, xmax = quant97.5, y = sim, color = type)) +
  # scale_x_log10() +
  facet_wrap(~n, ncol = 4) +
   theme_bw() +
  labs(x = "Posterior distribution", y = "Simulation",
       title = paste0("Posterior population distribution for HlyE_IgA/", i, 
                      " \nby number sampled and simulation")) +
  scale_color_manual(values = c("skyblue1", "dodgerblue4"))
)
}
fig_num <- fig_num +1

```


### RMSE analysis  
We will now calculate monte carlo bias comparing the median, 2.5% and 97.5% 
population-level posterior estimates for each parameter across samples taken. 
The estimates follow:  
$$
\mathrm{RMSE}_N =
\sqrt{
\frac{1}{S}
\sum_{s=1}^{S}
\left(
\hat{q}_{s,N} - q_{\mathrm{original}}
\right)^2
}
$$

Data preparation:
```{r error = F, warning = F, fig.height=12}

sim_post <- nepplot_pop_summ %>%
  filter(type == "Simulated posterior")

original_post <- nepplot_pop_summ %>%
  filter(type == "Original posterior") %>%
  dplyr::ungroup() %>%
  select(Parameter, Iso_type, median, quant2.5, quant97.5, type)

comparison <- sim_post %>%
  left_join(
    original_post %>%
      select(
        Parameter,
        Iso_type,
        original_lower = quant2.5,
        original_median = median,
        original_upper = quant97.5
      ),
    by = c("Parameter", "Iso_type")
  )

comparison <- comparison %>%
  mutate(
    error_lower = quant2.5 - original_lower,
    error_median = median - original_median,
    error_upper = quant97.5 - original_upper
  )

mc_bias <- comparison %>%
  group_by(n, Parameter, Iso_type) %>%
  summarise(
    bias_lower = mean(error_lower, na.rm = TRUE),
    bias_median = mean(error_median, na.rm = TRUE),
    bias_upper = mean(error_upper, na.rm = TRUE),
    .groups = "drop"
  )
mc_rmse <- comparison %>%
  group_by(n, Parameter, Iso_type) %>%
  summarise(
    rmse_lower = sqrt(mean(error_lower^2, na.rm = TRUE)),
    rmse_median = sqrt(mean(error_median^2, na.rm = TRUE)),
    rmse_upper = sqrt(mean(error_upper^2, na.rm = TRUE)),
    .groups = "drop"
  )

mc_rmse_long <- comparison %>%
  group_by(n, Parameter, Iso_type) %>%
  summarise(
    lower = sqrt(mean(error_lower^2, na.rm = TRUE)),
    median = sqrt(mean(error_median^2, na.rm = TRUE)),
    upper = sqrt(mean(error_upper^2, na.rm = TRUE)),
    .groups = "drop"
  ) %>%
  tidyr::pivot_longer(
    cols = c(lower, median, upper),
    names_to = "posterior_quantile",
    values_to = "RMSE"
  )


```

Figure `r fig_num`: Monte Carlo RMSE of median, 2.5% and 97.5% posterior distribution 
by number of simulated cases used for modeling and antigen isotype  
```{r error = F, warning = F, fig.height=12}

ggplot(
  mc_rmse_long,
  aes(x = n, y = RMSE, linetype = posterior_quantile,
      group = posterior_quantile)) +
  geom_line() +
  geom_point(size = 2) +
  facet_grid(
    Parameter ~ Iso_type,
    scales = "free_y"
  ) +
  theme_bw() +
  labs(
    x = "Number of subjects",
    y = "Monte Carlo RMSE",
    linetype = "Posterior summary"
  )+
    theme(legend.position = "top")

fig_num <- fig_num +1

```


### Modeled serokinetic curves  
Figure `r fig_num`: Serokinetic curves of simulated and original predictive 
posterior curves by number of simulations sampled and antigen isotype  
```{r error = F, warning = F, message = F}

# Need to also gather median from the original data model
nepal_pred_seroplot <- left_join(nepal_sees_output, nepal_sees_summ_bind,
                                by = "Parameter")

plot_pred_curves <- bind_rows(neppost_pred_sim, nepal_pred_seroplot) %>%
  mutate(Stratification = sim) %>%
  mutate(Stratification = ifelse(is.na(Stratification), "Original", 
                                 Stratification))
  
plot_list <- list()
for (i in unique(plot_pred_curves$n)) {
strat_levels <- unique(na.omit(plot_pred_curves$Stratification))
color_values <- setNames(
  ifelse(strat_levels == "Original", "red", "blue"),
  strat_levels
)
## Original data layer
plot_pred_curveplot <- plot_pred_curves %>%
  filter(n == i)
p_original <- plot_serocurve(
  plot_pred_curveplot %>% 
    filter(Stratification == "Original"),
  log_y = TRUE,
  show_ci = TRUE
) 
  # Make Original curve prominent
  p_original$layers <- lapply(
    p_original$layers,
    function(layer) {
      if (inherits(layer$geom, "GeomLine")) {
        layer$aes_params$color <- "red"
        layer$aes_params$alpha <- 1
        layer$aes_params$linewidth <- 1.2
      }
      layer
    }
  )
  
## Simulation layer
p_sim <- plot_serocurve(
  plot_pred_curveplot %>% 
    filter(Stratification != "Original"),
  log_y = TRUE,
  show_ci = FALSE
)
  p_sim$layers <- lapply(
    p_sim$layers,
    function(layer) {
      if (inherits(layer$geom, "GeomLine")) {
        layer$aes_params$color <- "blue"
        layer$aes_params$alpha <- 0.4
        layer$aes_params$linewidth <- 0.5
      }
      layer
    }
  )
  
plot_list[[i]] <- p_sim +
  p_original$layers +
  scale_color_manual(
    values = color_values,
    breaks = c(
      strat_levels[strat_levels != "Original"][1],
      "Original"
    ),
    labels = c("Simulation", "Original")
  ) +
  guides(
    color = guide_legend(
      title = NULL,
      override.aes = list(
        alpha = 1,
        linewidth = 1
      )
    ),
    fill = "none"
  ) +
  labs(title = paste0("n = ", i)) 
}
wrap_plots(plot_list, ncol = 2)  +
  plot_layout(guides = "collect") &
  theme(
    legend.position = "bottom"
  )


  
```


Figure `r fig_num`: Serokinetic curves of simulated and original population level 
posterior curves by number of simulations sampled and antigen isotype  
```{r error = F, warning = F, message = F}

# Need to also gather median from the original data model
nepal_pop_seroplot <- attributes(nepal_sees_output)$population_params 
nepal_pop_seroplot <- left_join(nepal_pop_seroplot, nepal_pop_summ_bind,
                                by = "Parameter")

plot_pop_curves <- bind_rows(nep_pop_data, nepal_pop_seroplot) %>%
  filter(Population_Parameter == "mu.par") %>%
  mutate(Stratification = sim) %>%
  mutate(Stratification = ifelse(is.na(Stratification), "Original", 
                                 Stratification)) %>%
  filter(is.na(n)==F)

plot_list <- list()
for (i in unique(plot_pop_curves$n)) {
strat_levels <- unique(na.omit(plot_pop_curves$Stratification))
color_values <- setNames(
  ifelse(strat_levels == "Original", "red", "blue"),
  strat_levels
)

# Setting up data first
plot_pop_curvesplot <- plot_pop_curves %>%
  filter(n == i)
attr(plot_pop_curvesplot, "population_params") <- plot_pop_curvesplot

## Original data layer
p_original <- plot_serocurve(
  plot_pop_curvesplot %>% 
    filter(Stratification == "Original"),
  log_y = TRUE,
  show_ci = TRUE,
  param_source = "population"
) 
  # Make Original curve prominent
  p_original$layers <- lapply(
    p_original$layers,
    function(layer) {
      if (inherits(layer$geom, "GeomLine")) {
        layer$aes_params$color <- "red"
        layer$aes_params$alpha <- 1
        layer$aes_params$linewidth <- 1.2
      }
      layer
    }
  )
  
## Simulation layer
p_sim <- plot_serocurve(
  plot_pop_curvesplot %>% 
    filter(Stratification != "Original"),
  log_y = TRUE,
  show_ci = FALSE,
  param_source = "population"
)
  p_sim$layers <- lapply(
    p_sim$layers,
    function(layer) {
      if (inherits(layer$geom, "GeomLine")) {
        layer$aes_params$color <- "blue"
        layer$aes_params$alpha <- 0.4
        layer$aes_params$linewidth <- 0.5
      }
      layer
    }
  )
  
plot_list[[i]] <- p_sim +
  p_original$layers +
  scale_color_manual(
    values = color_values,
    breaks = c(
      strat_levels[strat_levels != "Original"][1],
      "Original"
    ),
    labels = c("Simulation", "Original")
  ) +
  guides(
    color = guide_legend(
      title = NULL,
      override.aes = list(
        alpha = 1,
        linewidth = 1
      )
    ),
    fill = "none"
  ) +
  labs(title = paste0("n = ", i)) 
}
wrap_plots(plot_list, ncol = 2)  +
  plot_layout(guides = "collect") &
  theme(
    legend.position = "bottom"
  )

```
  

### Rhat values by number of samples
Rhat values by number of samples for predictive parameters
```{r results = 'hide', error = F, warning = F}

sim_pred_rhat <- neppost_pred_sim %>%
  filter(Subject == "newperson")
# Calculate Rhat
sim_pred_rhat <- sim_pred_rhat %>%
  group_by(Parameter, Iso_type, Stratification, sim, n) %>%
  group_modify(~{
    mat <- .x %>%
      arrange(Iteration, Chain) %>%
      pivot_wider(
        id_cols = Iteration,
        names_from = Chain,
        values_from = value
      ) %>%
      select(-Iteration) %>%
      as.matrix()
    tibble(Rhat = posterior::rhat(mat))
  }) %>%
  group_by(Parameter, Iso_type, n) %>%
  summarise(median = median(Rhat),
            lower = quantile(Rhat, 0.025),
            upper = quantile(Rhat, 0.975))

# sim_pred_rhat$n <- factor(sim_pred_rhat$n, c("3", "5", "10", "30", "50", 
#                                                          "100", "150", "188"))

ggplot(sim_pred_rhat, aes(x = as.numeric(n), y = median)) +
  geom_line(aes(x = as.numeric(n), y = median), stat = "identity") +
  geom_point(size = 2) +
  geom_linerange(aes(x = as.numeric(n), y = median, ymax = upper, ymin = lower)) +
  facet_grid(
    Parameter ~ Iso_type,
    scales = "free_y"
  ) +
  theme_bw() +
  labs(x = "Number of samples from posterior",
    y = "Rhat value"
  )+
    theme(legend.position = "top")

```

Rhat values by number of samples for population parameters
```{r results = 'hide', error = F, warning = F}

sim_pop_rhat <- nep_pop_data %>%
  filter(Population_Parameter == "mu.par")
# Calculate Rhat
sim_pop_rhat <- sim_pop_rhat %>%
  group_by(Parameter, Iso_type, Stratification, sim, n) %>%
  group_modify(~{
    mat <- .x %>%
      arrange(Iteration, Chain) %>%
      pivot_wider(
        id_cols = Iteration,
        names_from = Chain,
        values_from = value
      ) %>%
      select(-Iteration) %>%
      as.matrix()
    tibble(Rhat = posterior::rhat(mat))
  }) %>%
  group_by(Parameter, Iso_type, n) %>%
  summarise(median = median(Rhat),
            lower = quantile(Rhat, 0.025),
            upper = quantile(Rhat, 0.975))

# sim_pred_rhat$n <- factor(sim_pred_rhat$n, c("3", "5", "10", "30", "50", 
#                                                          "100", "150", "188"))

ggplot(sim_pop_rhat, aes(x = as.numeric(n), y = median)) +
  geom_line(aes(x = as.numeric(n), y = median), stat = "identity") +
  geom_point(size = 2) +
  geom_linerange(aes(x = as.numeric(n), y = median, ymax = upper, ymin = lower)) +
  facet_grid(
    Parameter ~ Iso_type,
    scales = "free_y"
  ) +
  theme_bw() +
  labs(x = "Number of samples from posterior",
    y = "Rhat value"
  )+
    theme(legend.position = "top")

```
