---
title: "Stan Backend: Multivariate Observation Model (Model B)"
format: html
---

```{r}
#| include: false
knitr::opts_chunk$set(
  collapse = TRUE,
  comment  = "#>",
  fig.width  = 7,
  fig.height = 5,
  eval = FALSE
)
```

## Overview

Model B extends the JAGS antibody kinetics model by replacing the
*K* independent univariate normal likelihoods (one per antigen-isotype) with a
single *K*-variate normal likelihood per time point.  This captures residual
correlations across antigen-isotype pairs within each observation, which the
independent model ignores.

The `run_mod_stan()` function implements Model B using RStan.  Its output is an
`sr_model` tibble with the same column layout as `run_mod()`, so all existing
plotting and summary functions (`plot_predicted_curve()`, `post_summ()`, etc.)
work without modification.

### Why Stan instead of JAGS?

- **LKJ prior** for correlation matrices — better than the Wishart prior in
  small samples (Gelman 2006).
- **NUTS sampler** — more efficient exploration of correlated posteriors.
- **Extendable** to Model C (Kronecker product covariance for cross-biomarker
  parameter correlations), which JAGS cannot handle.

### The two-phase kinetics curve

The mathematical model is identical to `model.jags`:

| Phase | Condition | Equation |
|-------|-----------|----------|
| Rising | $t \le t_1$ | $\log y(t) = \log y_0 + \beta t$ |
| Decay | $t > t_1$ | $\log y(t) = \frac{1}{1-\rho}\log\!\bigl[y_1^{1-\rho} - (1-\rho)\,\alpha\,(t - t_1)\bigr]$ |

where $\beta = (\log y_1 - \log y_0)/t_1$.

### Model B likelihood

$$
\log \mathbf{y}_{i,t} \sim \mathcal{N}_K\!\left(\boldsymbol{\mu}_{i,t},\; \Sigma_\varepsilon\right)
$$

$\Sigma_\varepsilon = \mathrm{diag}(\boldsymbol{\tau}_\varepsilon)\,\Omega_\varepsilon\,\mathrm{diag}(\boldsymbol{\tau}_\varepsilon)$,
with $\Omega_\varepsilon \sim \mathrm{LKJ}(2)$ and
$\tau_{\varepsilon,k} \sim \mathrm{Cauchy}^+(0, 2.5)$.

## Installation

Install `rstan` if you have not already:

```{r}
install.packages("rstan")
```

Install `serodynamics`:

```{r}
# install.packages("pak")
pak::pak("UCD-SERG/serodynamics")
```

## Quick start

```{r}
library(serodynamics)
library(dplyr)

# Simulate 20 subjects using typhoid curve parameters
set.seed(42)
sim_data <- serocalculator::typhoid_curves_nostrat_100 |>
  sim_case_data(n = 20)
```

### Fit Model B

```{r}
fit <- run_mod_stan(
  data   = sim_data,
  model  = "model_b",
  chains = 4,
  iter   = 2000,
  warmup = 1000,
  seed   = 1
)
```

### Inspect output

```{r}
# Same structure as run_mod() output
print(fit)

# Residual correlation matrix across antigen-isotype pairs
attr(fit, "Omega_eps")

# Residual covariance matrix
attr(fit, "Sigma_eps")
```

### Fitted and residual values

```{r}
attr(fit, "fitted_residuals") |> head()
```

### Visualize predicted curves

Because `fit` is an `sr_model` object, the standard plotting functions work
directly:

```{r}
plot_predicted_curve(fit, sim_data)
```

### Posterior summaries

```{r}
post_summ(fit)
```

## Comparing JAGS and Stan backends

Both `run_mod()` and `run_mod_stan()` return an `sr_model` tibble, so you can
compare them side-by-side:

```{r}
jags_fit <- run_mod(
  data   = sim_data,
  nchain = 4,
  nadapt = 500,
  nburn  = 500,
  nmc    = 500,
  niter  = 500
)

# Column names are identical
identical(names(jags_fit), names(fit))

# Compare posterior medians for y0
bind_rows(
  jags_fit  |> filter(Parameter == "y0") |> mutate(backend = "JAGS"),
  fit       |> filter(Parameter == "y0") |> mutate(backend = "Stan")
) |>
  group_by(backend, Iso_type) |>
  summarise(median_y0 = median(value), .groups = "drop")
```

## Data flow

```
as_case_data(raw)          # unchanged
    │
    ▼
prep_data(case_data,       # unchanged (add_newperson = FALSE used internally)
          add_newperson = FALSE)
    │
    ▼
prep_data_stan(prepped)    # NEW — reshape to Stan list
    │
    ▼
rstan::sampling(model_b)   # Stan NUTS sampler
    │
    ▼
postprocess_stan_output()  # NEW — stanfit → sr_model tibble
    │
    ▼
sr_model tibble            # identical structure to run_mod() output
```

## Session info

```{r}
sessioninfo::session_info()
```
